Visual Interrogation of Attention-Based Models for Natural Language Inference and Machine Comprehension

### 2018-06-01 EMNLP-demo-paper/result.tex
grammar fix
feedback from chris

%\begin{figure*}[t]
%\centering
%\vspace{-2mm}
% \includegraphics[width=1.0\linewidth]{NLI_MC_interface}
%  \vspace{-6mm}
% \caption{
%Illustration of different configurations for the natural language inference and machine comprehension tasks.
%}
%\label{fig:pipelineUpdate}
%\end{figure*}
\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{NLIexample}
  \vspace{-7mm}
 \caption{
An illustration of the attention editing process.
The dependency structure is shown in (a), where the two ``greens'' are decorating different nouns.
By removing the ``wrong'' alignment in (b), the original prediction \emph{entailment} is corrected to \emph{netural} in (c).
}
\vspace{-2mm}
\label{fig:NLIexample}
\end{figure*}

@@

\subsection{Natural Language Inference}
\label{sec:NLIexample}
The NLI task predicts the relationship between a premise sentence (P) and a hypothesis sentence (H).
The attention matrix captures the alignment of words between these two sentences.
Here we give an example of how a wrong prediction can be correct by editing the attention value.

@@

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pair (P:``A woman in a green jacket is drinking tea.'' H:``A woman is drinking green tea.'') is predicted to be \emph{entailment}, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two \emph{greens} decorate different nouns, which potentially leads to the wrong prediction. The grammar structure is visually presented by dependency tree.
%i.e., \emph{green} in \textbf{P} is attached to ``jacket''.
However, the model does not has access to the grammar information and mistakenly assumes the two \emph{greens} are describing the same thing.
% (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two "greens" (see (c)(b)).
As expected, the prediction label is corrected (\emph{neutral}).

@@

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{MC_interface}
  \vspace{-7mm}
 \caption{
In the machine comprehension visualization interface (a), the $p1$, $p2$ colored bar (in $a3$) illustrates the predicted start and end index of the answer in the context (the deeper the red, the higher the probability). The most likely answer is shown in ($a1$). The global and local attention are visualized by ($a2$, $a3$).
%
We can evaluate the robustness of the prediction by perturbing the question sentence ($b1$, $b3$). As illustrated in ($b1$, $b2$) by removing the word ``partial'', the model still finds the correct answer (albeit different, as the sentence perturbation changes the exact meaning of the question). 
}
\label{fig:MCexample}
\end{figure*}

As illustrated in Figure~\ref{fig:MCexample}, we represent \emph{att2} as colored bars with \emph{yellow-green-blue} colormap. Each rectangular bar corresponds to one sentence. The user can focus on individual sentences by clicking on the rectangular bar (see Figure~\ref{fig:MCexample}).
%The proposed visualization help reveal the potential alignment issues in the machine comprehension model.
The $p1$, $p2$ colored bars (\emph{white-red} colormap) illustrate the predicted probabilities of the start and the end index for the answer (the deeper the higher). %The span of text defined by the beginning and end index with the highest probability is the answer.
%
%In Figure~\ref{fig:MCexample} ($a3$), there are two spans of text that correspond to higher probabilities. 
In Figure~\ref{fig:MCexample}(a3), the sentence containing the answer exhibits good alignment with the question (e.g. ``Two'' with ``many'').
%In (a3), we examine the sentence that contains the answer, which exhibits excellent alignment with the question (e.g., ``Two'' is aligned with ``many''). 
%The spans of text with high probability are ``Two'' and ``No'' (highlight in orange), both align well with the word ``many''.
Interestingly, the number ``9'' (in ``Falcon 9') is also aligned with ``many'', which may potentially lead to problems.

The user can explore the robustness of the model by examining how the prediction varies when the question is perturbed.
%
As illustrated in ($b1$, and $b2$), the perturbation removes the word ``partial'' in the original sentence, which leads the model to produce a different yet correct answer (``No''). Referring back to ($a1$), we can see the word ``No'' exhibits the second highest probability for the original question.
%
The user can also manually edit the text. As shown in ($b3$). The model still produces the correct answer when changing the question from ``how many'' to ``which''.
%One possible explanation is that the word ``Two'' has a much higher \emph{att2} attention value compared to ``9'', which may contribute to the current outcome.

>>>>

%\begin{figure*}[t]
%\centering
%\vspace{-2mm}
% \includegraphics[width=1.0\linewidth]{NLI_MC_interface}
%  \vspace{-6mm}
% \caption{
%Illustration of different configurations for the natural language inference and machine comprehension tasks.
%}
%\label{fig:pipelineUpdate}
%\end{figure*}
\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{NLIexample}
  \vspace{-7mm}
 \caption{
An illustration of the attention editing process.
The dependency structure is shown in (a), where the two ``greens'' decorate different nouns.
By removing the ``wrong'' alignment in (b), the original prediction \emph{entailment} is corrected to \emph{neutral} in (c).
}
\vspace{-2mm}
\label{fig:NLIexample}
\end{figure*}

@@

\subsection{Natural Language Inference}
\label{sec:NLIexample}
The NLI task predicts the relationship between a premise sentence (P) and a hypothesis sentence (H).
The attention matrix captures the alignment of words between these two sentences.
Here we give an example of how a wrong prediction can be corrected by editing the attention value.

@@

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pair (P:``A woman in a green jacket is drinking tea.'' H:``A woman is drinking green tea.'') is predicted to be \emph{entailment}, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two \emph{greens} decorate different nouns, which potentially leads to the wrong prediction. The grammar structure is visually presented by a dependency tree.
%i.e., \emph{green} in \textbf{P} is attached to ``jacket''.
However, the model does not have access to the grammar information and mistakenly assumes the two \emph{greens} are describing the same thing.
% (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two "greens" (see (c)(b)).
As expected, the prediction label is corrected (\emph{neutral}).

@@

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{MC_interface}
  \vspace{-7mm}
 \caption{
In the machine comprehension visualization interface (a), the $p1$, $p2$ colored bar (in $a3$) illustrates the predicted start and end index of the answer in the context (the deeper the red, the higher the probability). The most likely answer is shown in ($a1$). The global attention and local attention are visualized by ($a2$, $a3$).
%
We can evaluate the robustness of the prediction by perturbing the question sentence ($b1$, $b3$). As illustrated in ($b1$, $b2$), by removing the word ``partial'', the model still finds the correct answer (albeit different, as the sentence perturbation changes the exact meaning of the question). 
}
\label{fig:MCexample}
\end{figure*}

As illustrated in Figure~\ref{fig:MCexample}, we represent \emph{att2} as colored bars with a \emph{yellow-green-blue} colormap. Each rectangular bar corresponds to one sentence. The user can focus on individual sentences by clicking on the rectangular bar (see Figure~\ref{fig:MCexample}).
%The proposed visualization help reveal the potential alignment issues in the machine comprehension model.
The $p1$, $p2$ colored bars (\emph{white-red} colormap) illustrate the predicted probabilities of the start and the end index for the answer (the deeper the higher). %The span of text defined by the beginning and end index with the highest probability is the answer.
%
%In Figure~\ref{fig:MCexample} ($a3$), there are two spans of text that correspond to higher probabilities. 
In Figure~\ref{fig:MCexample}(a3), the sentence containing the answer exhibits good alignment with the question (e.g., ``Two'' with ``many'').
%In (a3), we examine the sentence that contains the answer, which exhibits excellent alignment with the question (e.g., ``Two'' is aligned with ``many''). 
%The spans of text with high probability are ``Two'' and ``No'' (highlight in orange), both align well with the word ``many''.
Interestingly, the number ``9'' (in ``Falcon 9') is also aligned with ``many'', which may lead to problems.

The user can explore the robustness of the model by examining how the prediction varies when the question is perturbed.
%
As illustrated in ($b1$, and $b2$), the perturbation removes the word ``partial'' in the original sentence, which leads the model to produce a different yet correct answer (``No''). Referring to ($a1$), we can see the word ``No'' exhibits the second highest probability for the original question.
%
The user can also manually edit the text. As shown in ($b3$), the model still produces the correct answer when changing the question from ``how many'' to ``which''.
%One possible explanation is that the word ``Two'' has a much higher \emph{att2} attention value compared to ``9'', which may contribute to the current outcome.
